---
title: Frontiers in Difference-in-Differences
subtitle: Alternative Identification Strategies
format: clean-revealjs
author:
  - name: Brantly Callaway
    orcid: 0000-0000-0000-0000
    email: brantly.callaway@uga.edu
    affiliations: University of Georgia
date: last-modified
---

## Introduction

```{r echo=FALSE}
# load packages
library(revealEquations)
```

$\newcommand{\E}{\mathbb{E}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\var}{\mathrm{var}}
\newcommand{\cov}{\mathrm{cov}}
\newcommand{\Var}{\mathrm{var}}
\newcommand{\Cov}{\mathrm{cov}}
\newcommand{\Corr}{\mathrm{corr}}
\newcommand{\corr}{\mathrm{corr}}
\newcommand{\L}{\mathrm{L}}
\renewcommand{\P}{\mathrm{P}}
\newcommand{\independent}{{\perp\!\!\!\perp}}
\newcommand{\indicator}[1]{ \mathbf{1}\{#1\} }$
<span class="alert">Recap:</span>

* We have been following the high-level strategy of (1) targeting disaggregated parameters and then (2) combining them.  

* Part 4: allow for more complicated treatment regimes, but use difference-in-differences

* Part 5: go back to staggered treatment setting, but use different identification assumptions

<span class="alert">Examples in this part</span>

1. Change-in-Changes

2. Interactive Fixed Effects



## Introduction to Change-in-Changes

The idea of change-in-changes comes from Athey and Imbens (2006) and builds on work on estimating non-separable production function models.  They consider the case where
\begin{align*}
  Y_{it}(0) = h_t(U_{it})
\end{align*}
where $h_t$ is a nonparametric, time-varying function.  To me, it is helpful to think of $U_{it} = \eta_i + e_{it}$.  This model (for the moment) generalizes the model that we used to rationalize parallel trends: $Y_{it}(0) = \theta_t + \eta_i + e_{it}$.

. . .

<span class="alert">Additional Conditions:</span>

1. $U_{it} \overset{d}{=} U_{it'} | G$.  In words: the distribution of $U_{it}$ does not change over time given a particular group.  However, the distribution of $U_{it}$ can vary across groups.

2. $U_{it}$ is scalar

3. $h_t$ is stictly monotonically increasing $\implies$ we can invert it.

4. Support condition: $\mathcal{U}_g \subseteq \mathcal{U}_0$ (support of $U_{it}$ for the treated group is a subset of the support of $U_{it}$ for the untreated group) <!--$\implies \textrm{support}(Y_{it}(0))$ for the treated group is a subset of the-->

<!-- 1. is more restrictive, 2. is arguably similar, 3. is similar, 4. is not require for DID can extrapolate -->



## Change-in-Changes Identification

Under the conditions described above, you can show that

\begin{align*}
  ATT(g,t) = \E[Y_t | G=g] - \E\Big[Q_{Y_t(0)|U=1}\big(F_{Y_{g-1}(0)|U=1}(Y_{g-1}(0))\big) | G=g \Big]
\end{align*}
where $Q_{Y_t(0)|U=1}(\tau)$ is the $\tau$-th quantile of $Y_t(0)$ for the never-treated group (e.g., if $\tau=0.5$, it is the median of $Y_t(0)$ for the never-treated group).

. . .

* [As an interesting side-comment, this is derived in Athey and Imbens (2006), way before recent work on group-time average treatment effects, and it is pretty much exactly analogous to the "first step" that we have been emphasizing]



## Intuition for Change-in-Changes

<span class="alert">Intuition: </span> Notice that, under parallel trends, we can re-write
\begin{align*}
  ATT(g,t) = \E[Y_t|G=g] - \E\left[ \Big(\E[Y_t | U=1] - \E[Y_{g-1} | U=1]\Big) + Y_{g-1} | G=g \right]
\end{align*}
which we can think of as: compare observed outcomes to, (an average of) taking observed outcomes in the pre-treatment period and accounting for how outcomes change over time in the untreated group across the same periods

. . .

For CIC, the intuition is the same, except the way that we "account for" how outcomes change over time during the same periods for the untreated group is a different.

. . .

Because these are different transformations, DID and CIC are non-nested approaches.



## Comments

CIC is a nice approach in many applications 

* In addition, to recovering $ATT(g,t)$, it is also possible to recover <span class="highlight">quantile treatment effect parameters</span> in this setting (these can allow you to more effectively study treatment effect heterogeneity and are closely related to social welfare calculatations/comparisons)

. . .

Though it is less commonly used in empirical work than DID.

* Need to estimate quantiles

* Harder to include covariates (due to needing to estimate quantiles).  I think (not 100% sure though) that it is not possible (at least not obvious) if one can do a doubly robust version of CIC.

* Support conditions can have real bite in some applications

* Not as much software support



## Minimum Wage Application

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(did)
library(BMisc)
library(twfeweights)
library(fixest)
library(modelsummary)
library(ggplot2)
library(pte)
library(qte)
library(ife)
library(dplyr)
library(tidyr)
load("apm.RData")
```

```{r eval=FALSE, message=FALSE}
## change-in-changes
data2$G2 <- data2$G
cic_res <- qte::cic2(yname="lemp", 
                     gname="G2",
                     tname="year",
                     idname="id",
                     data=data2,
                     boot_type="empirical",
                     cl=4)
ggpte(cic_res)
```

$\widehat{ATT}^O = -0.059$, $\textrm{s.e}(\widehat{ATT}^O) = 0.009$.  (This is very close to our estimate using DID before: $-0.057$)



## Minimum Wage Application

```{r, echo=FALSE, cache=TRUE, fig.align="center", fig.width=14, fig.height=8, message=FALSE, warning=FALSE}
# change-in-changes
data2$G2 <- data2$G
cic_res <- qte::cic2(yname="lemp", 
                     gname="G2",
                     tname="year",
                     idname="id",
                     data=data2,
                     boot_type="empirical",
                     cl=4)
ggpte(cic_res) + geom_hline(yintercept=0, size=1.2)
```



## Interactive Fixed Effects

Earlier we discussed this model for untreated potential outcomes $Y_{it}(0) = h_t(\eta_i, e_{it})$ and argued that it was too general to make much progress on.  

. . .

An intermediate case is an interactive fixed effects model for untreated potential outcomes:
\begin{align*}
  Y_{it}(0) = \theta_t + \eta_i + \lambda_i F_t + e_{it}
\end{align*}

* $\lambda_i$ is often referred to as "factor loading" (notation above implies that this is a scalar, but you can allow for higher dimension)

* $F_t$ is often referred to as a "factor"

* $e_{it}$ is idioyncratic in the sense that $\E[e_{it} | G_i=g] = 0$ for all groups

. . .

In our context, though, it makes sense to interpret these as

* $\lambda_i$ unobserved heterogeneity (e.g., individual's unobserved skill)

* $F_t$ the time-varying "return" unobserved heterogeneity (e.g., return to skill) 



## Interactive Fixed Effects

Interactive fixed effects models for untreated potential outcomes generalize some other important cases:

. . .

<span class="alert">Example 1: </span> Suppose we observe $\lambda_i$, then this amounts to the regression adjustment version of DID with a time-invariant covariate considered earlier

. . .

<span class="alert">Example 2: </span> Suppose you know that $F_t = t$, then this leads to a *unit-specific linear trend model*:
\begin{align*}
  Y_{it}(0) = \theta_t + \eta_i + \lambda_i t + e_{it}
\end{align*}

. . .

To allow for $F_t$ to change arbitrarily over time is harder...

. . .


<span class="alert">Example 3: </span> Interactive fixed effects models also provide a connection to "large-T" approaches such as synthetic control and synthetic DID (Abadie, Diamond, and Hainmueller (2010), Arkhangelsky et al. (2021))

* e.g., one of the motivations of the SCM in ADH-2010 is that (given large-T) constructing a synthetic control can balance the factor loadings in an interactive fixed effects model for untreated potential outcomes



## Interactive Fixed Effects

Interactive fixed effects models allow for violations of parallel trends:

\begin{align*}
  \E[\Delta Y_{it}(0) | G=g] = \Delta \theta_t + \E[\lambda_i|G=g]\Delta F_t
\end{align*}
which can vary across groups.

Example: If $\lambda_i$ is "ability" and $F_t$ is increasing over time, then (even in the absence of the treatment) groups with higher mean "ability" will tend to increase outcomes more over time than less skilled groups



## How can you recover $ATT(g,t)$ here?

There are a lot of ideas.  Probably the most prominent idea is to directly estimate the model for untreated potential outcomes and impute

* See Xu (2017) and Gobillon and Magnac (2018) for substantial detail on this front

* For example, Xu (2017) uses Bai (2009) principal components approach to estimate the model.  This is a bit different in spirit from what we have been doing before as this argument requires the number of time periods to be "large"



## Alternative Approaches with Fixed-T

<span class="alert">Very Simple Case:</span>

* $\mathcal{T}=4$

* 3 groups: 3, 4, $\infty$

* We will target $ATT(3,3) = \E[\Delta Y_{i3} | G_i=3] - \underbrace{\E[\Delta Y_{i3}(0) | G_i=3]}_{\textrm{have to figure out}}$

. . .

In this case, given the IFE model for untreated potential outcomes, we have:
\begin{align*}
  \Delta Y_{i3}(0) &= \Delta \theta_3 + \lambda_i \Delta F_3 + \Delta e_{i3} \\
  \Delta Y_{i2}(0) &= \Delta \theta_2 + \lambda_i \Delta F_3 + \Delta e_{i2} \\
\end{align*}

. . .

The last equation implies that
\begin{align*}
  \lambda_i = \Delta F_2^{-1}\Big( \Delta Y_{i2}(0) - \Delta \theta_2 - \Delta e_{i2} \Big)
\end{align*}
Plugging this back into the first equation (and combining terms), we have $\rightarrow$



## Alternative Approaches with Fixed-T

From last slide, combining terms we have that

\begin{align*}
  \Delta Y_{i3}(0) = \underbrace{\Big(\Delta \theta_3 - \frac{\Delta F_3}{\Delta F_2} \Delta \theta_2 \Big)}_{=: \theta_3^*} + \underbrace{\frac{\Delta F_3}{\Delta F_2}}_{=: F_3^*} \Delta Y_{i2}(0) + \underbrace{\Delta e_{i3} - \frac{\Delta F_3}{\Delta F_2} \Delta e_{i2}}_{=: v_{i3}}
\end{align*}

. . .

Now (momentarily) suppose that we (somehow) know $\theta_3^*$ and $F_3^*$.  Then,

\begin{align*}
  \E[\Delta Y_{i3}(0) | G_i=3] = \theta_3^* + F_3^* \underbrace{\E[\Delta Y_{i2}(0) | G_i = 3]}_{\textrm{identified}} + \underbrace{\E[v_{i3}|G_i=3]}_{=0}
\end{align*}

$\implies$ this term is identified; hence, we can recover $ATT(3,3)$.



## Alternative Approaches with Fixed-T

From last slide, combining terms we have that

\begin{align*}
  \Delta Y_{i3}(0) = \underbrace{\Big(\Delta \theta_3 - \frac{\Delta F_3}{\Delta F_2} \Delta \theta_2 \Big)}_{=: \theta_3^*} + \underbrace{\frac{\Delta F_3}{\Delta F_2}}_{=: F_3^*} \Delta Y_{i2}(0) + \underbrace{\Delta e_{i3} - \frac{\Delta F_3}{\Delta F_2} \Delta e_{i2}}_{=: v_{i3}}
\end{align*}

<span class="alert">How can we recover $\theta_3^*$ and $F_3^*$?</span>

. . .

Notice: this involves untreated potential outcomes through period 3, and we have groups 4 and $\infty$ for which we observe these untreated potential outcomes.  This suggests using those groups.

* However, this is not so simple because, by construction, $\Delta Y_{i2}(0)$ is correlated with $v_{i3}$ (note: $v_{i3}$ contains $\Delta e_{i2} \implies$ they will be correlated by construction)

* We need some exogenous variation (IV) to recover the parameters $\rightarrow$



## Alternative Approaches with Fixed-T

There are a number of different ideas here:

. . .

* Make additional assumptions ruling out serial correlation in $e_{it}$ $\implies$ can use lags of outcomes as instruments:

    * But this is seen as a strong assumption in many applications (Bertrand, Duflo, Mullainathan (2004))
    
. . .
    
* Alternatively can introduce covariates and make auxiliary assumptions about them (Callaway and Karami (2023) and Brown, Butts, and Westerlund (2023))

. . .
    
* However, it turns out that, with staggered treatment adoption, you can recover $ATT(3,3)$ essentially for free (Callaway and Tsyawo (2023)).



## Alternative Approaches with Fixed-T

In particular, notice that, given that we have two distinct untreated groups in period 3: group 4 and group $\infty$, then we have two moment conditions:

\begin{align*}
  \E[\Delta Y_{i3}(0) | G=4] &= \theta_3^* + F_3^* \E[\Delta Y_{i2}(0) | G=4] \\
  \E[\Delta Y_{i3}(0) | G=\infty] &= \theta_3^* + F_3^* \E[\Delta Y_{i2}(0) | G=\infty] \\
\end{align*}
We can solve these for $\theta_3^*$ and $F_3^*$, then use these to recover $ATT(3,3)$.

. . .

* The main requirement is that $\E[\lambda_i | G=4] \neq \E[\lambda_i|G=\infty]$ (relevance condition)

* Can scale this argument up for more periods, groups, and IFEs

* Relative to other approaches, the main drawback is that can't recover as many $ATT(g,t)$'s; e.g., in this example, we can't recover $ATT(3,4)$ or $ATT(4,4)$ which might be recoverable in other settings



## Minimum Wage Application

```{r eval=FALSE, message=FALSE}
# staggered ife
data4 <- subset(data3, G %in% c(2007,2006,0))
sife_res <- ife::staggered_ife2(yname="lemp",
                                gname="G",
                                tname="year",
                                idname="id",
                                data=data4,
                                nife=1)

did::ggdid(sife_res$att_gt)
```



## Minimum Wage Application

```{r, echo=FALSE, cache=TRUE, fig.align="center", fig.width=14, fig.height=8, message=FALSE, warning=FALSE}

# staggered ife
data4 <- subset(data3, G %in% c(2007,2006,0))
sife_res <- ife::staggered_ife2(yname="lemp",
                                gname="G",
                                tname="year",
                                idname="id",
                                data=data4,
                                nife=1)

did::ggdid(sife_res$att_gt)
```



## Summary

This section has emphasized alternative approaches to DID and LO to recover disaggregated treatment effect parameters:

* Change-in-Changes

* Interactive fixed effects models  

We have targeted $ATT(g,t)$.  Moving to more aggregated treatment effect parameters such as $ATT^{ES}(e)$ or $ATT^O$ is the same as before.



## Summary

I want to emphasize the high-level thought process one last time for using/inventing heterogeneity robust causal inference procedures with panel data: <!--some off-script application:-->

* Step 1: target disaggregated parameters directly using whatever approach you think would work well for recovering the $ATT$ for a fixed "group" and "time"

* Step 2: if desired, combine those disaggregated parameters into lower dimensional parameter that you may be able to estimate better and report more easily; hopefully, you can provide some motivation for this aggregated parameter




## Conclusion

<span class="highlight">Thank you</span> very much for having me!  

<br>

<span class="alert">Contact Information: </span>brantly.callaway@uga.edu

<span class="alert">Code and Slides: </span> [Available here](https://github.com/Mixtape-Sessions/Frontiers-in-DID)


<span class="alert">Papers:</span>

* Callaway (2023, *Handbook of Labor, Human Resources and Population Economics*), [[[published version](https://link.springer.com/referenceworkentry/10.1007/978-3-319-57365-6_352-1)]] &nbsp; [[[draft version](https://bcallaway11.github.io/files/Callaway-Chapter-2022/main.pdf)]]; the draft version is ungated and very similar to the published version.

* Today is also based on the not-yet-made-publicly available manuscript Baker, Callaway, Cunningham, Goodman-Bacon, and Sant'Anna (be on the lookout for it over the next few days)


