---
title: Frontiers in Difference-in-Differences
subtitle: Introduction to Panel Data Approaches to Causal Inference
format: clean-revealjs
author:
  - name: Brantly Callaway
email: brantly.callaway@uga.edu
affiliations: University of Georgia
date: October 17, 2023
knitr:
  opts_chunk:
    echo: true
---

```{r echo=FALSE}
library(revealEquations)
```

## Introduction  

$\newcommand{\E}{\mathbb{E}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\var}{\mathrm{var}}
\newcommand{\cov}{\mathrm{cov}}
\newcommand{\Var}{\mathrm{var}}
\newcommand{\Cov}{\mathrm{cov}}
\newcommand{\Corr}{\mathrm{corr}}
\newcommand{\corr}{\mathrm{corr}}
\newcommand{\L}{\mathrm{L}}
\renewcommand{\P}{\mathrm{P}}
\newcommand{\independent}{{\perp\!\!\!\perp}}
\newcommand{\indicator}[1]{ \mathbf{1}\{#1\} }$
Panel data gives researchers the opportunity to follow the same person, firm, location, etc. over multiple time periods

. . .

Having this sort of data seems fundamentally useful for learning about causal effects of some treatment/policy variable.

. . .

To see this, the <span class="alert">fundamental problem of causal inference</span> is that we can either see a unit's treated or untreated potential outcomes (but not both)

. . .

However, with panel data, this is not 100% true.  In some cases, we can see both a unit's treated and untreated potential outcome outcome...just at different points in time

* Example: 2 time periods, a unit is untreated in the first period but becomes treated in the second period

. . .

This seems extremely useful for learning about causal effects



## Introduction

Panel data approaches are also extremely common in empirical work

* Currie, Kleven, and Zwiers (AER P&P, 2020): 25% of NBER working papers in applied micro use difference-in-differences (i.e., a subset of panel data approaches to causal inference)

. . .

Some of this is likely due to the reasons mentioned above, but (at least as importantly) its popularity is due to wide availability of panel data


<span class="alert-blue">References: </span> 

  * Callaway (2023, *Handbook of Labor, Human Resources and Population Economics*)
  
  * Baker, Callaway, Cunningham, Goodman-Bacon, Sant'Anna (2023, draft is days away)


## Overview of Different Approaches to Causal Inference with Panel Data

<span class="alert">Types of Panel Data Approaches to Causal Inference:</span>

* Difference-in-differences

* Conditioning on lagged outcomes

* Unit-specific linear trends, interactive fixed effects models, change-in-changes, triple differences, others...




## Setup

<span class="alert">Setting: Exploit a data structure where the researcher observes:</span>

1. Multiple periods of data

2. Some pre-treatment data for all units

3. Some units become treated while other units remain untreated

. . .

(In my view) this particular data setup is a key distinguishing feature of the approaches that we'll mainly talk about today relative to traditional panel data models (i.e., fixed effects, dynamic panel, etc.)

* This setup also explains why the methods we consider today are often grouped among "natural experiment" types of methods such as IV or RD.

. . .

<span class="alert-blue">Running Examples:</span> 

* Causal effects of a state-level minimum wage increase on employment

* Causal effects of job displacement



## Setup


Modern approaches also typically allow for <span class="alert">treatment effect heterogeneity</span>

* That is, that effects of the treatment can vary across different units in potentially complicated ways


This is going to be a major issue in the discussion below

We'll consider implications for various estimation strategies as well as for "traditional" regression approaches



## Difference-in-differences

There have been a number of notable advances in recent years on panel data approaches to causal inference.  

* Many of these have been in the context of differences-in-differences

* That said, a number of the insights in this literature extend to other identification strategies


## The Logic of DID

<span class="alert-blue">Intuition for DID identification strategy</span> is to compare:

- The change in outcomes over time for units that participate in the treatment to
    
- The change in outcomes over time for units that didn't participate in the treatment 

. . .

Rough explanation: Compares a treated unit's outcomes to its past outcomes while making adjustment for <span class="alert">common shocks</span> using the comparison group.  


## Textbook Case with Two Periods

<span class="alert">Data: </span> 

* 2 periods: $t^*$, $t^*-1$ 
    * No one treated until period $t^*$
    * Some units remain untreated in period $t^*$

* 2 groups: $D=1$ or $D=0$ (treated and untreated)

. . .

<span class="alert">Potential Outcomes: </span> $Y_{it}(1)$ and $Y_{it}(0)$

. . .

<span class="alert">Observed Outcomes: </span> $Y_{it^*}$ and $Y_{it^*-1}$

\begin{align*}
  Y_{it^*} = D_i Y_{it^*}(1) +(1-D_i)Y_{it^*}(0) \quad \textrm{and} \quad Y_{it^*-1} = Y_{it^*-1}(0)
\end{align*}


## Target Parameter

<span class="alert">Average Treatment Effect on the Treated: </span>
$$ATT = \E[Y_{t^*}(1) - Y_{t^*}(0) | D=1]$$

Explanation: Mean difference between treated and untreated potential outcomes in the second period among the treated group


```{r echo=FALSE, results="asis"}

title <- "DID with Two Periods"

before <- "<span class=\"alert\">Parallel Trends Assumption: </span>
$$\\E[\\Delta Y_{t^*}(0) | D=1] = \\E[\\Delta Y_{t^*}(0) | D=0]$$
Explanation: Mean path of untreated potential outcomes is the same for the treated group as for the untreated group

. . .

<span class=\"alert\">Identification: </span>Under PTA, we can identify $ATT$:"

after <- "

. . .

$\\implies ATT$ is identified can be recovered by the difference in outcomes over time (difference 1) relative to the difference in outcomes over time for the untreated group (difference 2)"

eqlist <- list("ATT &= \\E[\\Delta Y_{t^*} | D=1] - \\E[\\Delta Y_{t^*}(0) | D=1]", 
               "&= \\E[\\Delta Y_{t^*} | D=1] - \\E[\\Delta Y_{t^*} | D=0]")

step_by_step_eq(eqlist=eqlist,
                before=before,
                after=after,
                title=title)
```


## TWFE Regression

TWFE regression: $$Y_{it} = \theta_t + \eta_i + \alpha D_{it} + e_{it}$$

<span class="alert-blue">Some more things to point out:</span>

. . .

- Even though it looks like this model has restricted the effect of participating in the treatment to be constant (and equal to $\alpha$) across all individuals,

    TWFE (in this case) is actually <span class="alert">robust</span> to treatment effect heterogeneity. To see this, notice that (with two periods) the previous regression is equivalent to
\begin{align*}
  \Delta Y_{it} = \Delta \theta_t + \alpha \Delta D_{it} + \Delta e_{it}
\end{align*}
This is fully saturated in $\Delta D_{it}$ (which is binary) $\implies$
\begin{align*}
  \alpha = \E[\Delta Y_{it}|D_{it}=1] - \E[\Delta Y_{it}|D=0] = ATT
\end{align*}
    




## TWFE Regression
    
It's easy to make the TWFE regression more complicated:

  - Multiple time periods
  
  - Variation in treatment timing
  
  - More complicated treatments
  
  - Introducing additional covariates


Unfortunately, the robustness of TWFE regressions to treatment effect heterogeneity or these more complicated (and empirically relevant) settings does not seem to hold

. . .

- Much of the recent (mostly negative) literature on TWFE in the context of DID has considered these types of "realistic" settings


## More Complicated Treatment Regimes

The arguments above are fairly easy and well-known.

. . .

Most applications, however, involve more complicated settings (more periods, more complicated treatment regimes, etc.)

. . .

One of the most active areas in causal inference with panel data in the past few years has been to these more "realistic" settings

. . .

A lot of these advancements have been in a DID framework (so I will emphasize this below)

* However, I think that a lot of the same insights apply to other identification strategies as well

* I'll make the case that you can just "substitute in", say, LO unconfoundedness in the "first step" for parallel trends, and a lot of the same arguments go through

* If we have time, I'll argue that you can use other identification strategies in the "first step" such as interactive fixed effects models and change-in-changes

. . .

The first more complicated treatment regime that we'll discuss is <span class="alert">staggered treatment adoption</span>



## Setup w/ Staggered Treatment Adoption

- $\mathcal{T}$ time periods

. . .

- Units can become treated at different points in time
    - <span class="alert-blue">staggered treatment adoption: </span> Once a unit becomes treated they remain treated.
    - $D_{it}$ &mdash; treatment indicator.  In math, staggered treatment adoption means: $D_{it-1}=1 \implies D_{it}=1$.
    - $G_i$ &mdash; a unit's <span class="alert-blue">group</span> &mdash; the time period that unit becomes treated.  Also, define $U_i=1$ for never-treated units and $U_i=0$ otherwise.

. . .

<span class="alert">Examples:</span>

* Government policies that roll out in different locations at different times (minimum wage is close to this over short time horizons)

* "Scarring" treatments: e.g., job displacement does not typically happen year after year, but rather labor economists think of being displaced as changing a person's "state" (the treatment is more like: has a person ever been displaced) 



## Setup w/ Staggered Treatment Adoption

- Potential outcomes: $Y_{it}(g)$ - the outcome that unit $i$ would experience in time period $t$ if they became treated in period $g$.

. . .

- Untreated potential outcome: $Y_{it}(0)$ - the outcome unit $i$ would experience in time period $t$ if they did not participate in the treatment in any period.  

. . .

- Observed outcome: $Y_{it}=Y_{it}(G_i)$

. . .

- No anticipation condition: $Y_{it} = Y_{it}(0)$ for all $t < G_i$ (pre-treatment periods for unit $i$)



## Unit-Level Treatment Effects

<span class="alert">Unit-level treatment effect</span>
$$\tau_{it}(g) = Y_{it}(g) - Y_{it}(0)$$

. . .

<span class="alert">Average treatment effect for unit $i$</span> (across time periods):
$$\bar{\tau}_i(g) = \frac{1}{\mathcal{T} - g + 1} \sum_{t=g}^{\mathcal{T}} \tau_{it}(g)$$



## Target Parameters

* <span class="alert">Group-time average treatment effects</span> 
\begin{align*}
  ATT(g,t) = \E[ \tau_{it}(G) | G=g]
\end{align*}
Explanation: $ATT$ for group $g$ in time period $t$

. . .

* <span class="alert">Event Study</span> 
\begin{align*}
  ATT^{ES}(e) = \E[\tau_{i,g+e}(G) | G \in \mathcal{G}_e]
\end{align*}
where $\mathcal{G}_e$ is the set of groups observed to have experienced the treatment for $e$ periods at some point.

    Explanation: $ATT$ when units have been treated for $e$ periods

<!--In math: $\mathcal{G}_e = \{g : (g+e) \in [2,T] \textrm{ and } g > 0\}$-->

. . .

* <span class="alert">Overall ATT</span> 
\begin{align*}
  ATT^O = \E[\bar{\tau}_i(G) | U=0]
\end{align*}
Explanation: $ATT$ across all units that every participate in the treatment



## Target Parameters

To understand the discussion later, it is also helpful to think of $ATT(g,t)$ as a <span class="alert">building block</span> for the other parameters discussed above.

. . .

Notice that:

\begin{align*}
  ATT^{ES}(e) = \sum_{g \in \bar{\mathcal{G}}} w^{ES}(g,e) ATT(g,g+e) \qquad \textrm{ and } \qquad ATT^O = \sum_{g \in \bar{\mathcal{G}}} \sum_{t=g}^{\mathcal{T}} w^O(g,t) ATT(g,t)
\end{align*}
where
\begin{align*}
  w^{ES}(g,e) = \indicator{g \in \mathcal{G}_e} \P(G=g|G\in \mathcal{G}_e) \qquad \textrm{and} \qquad
w^O(g,t) = \frac{\P(G=g|U=0)}{\mathcal{T}-g+1}
\end{align*}

. . .

<br>

In other words, if we can identify/recover $ATT(g,t)$, then we can proceed to recover $ATT^{ES}(e)$ and $ATT^O$.



## DID Identification of $ATT(g,t)$

<br>

::: {.callout-note}

## Multiple Period Version of Parallel Trends Assumption

For all groups $g \in \bar{\mathcal{G}}$ (all groups except the never-treated group) and for all time periods $t=2,\ldots,\mathcal{T}$,
\begin{align*}
  \E[\Delta Y_{t}(0) | G=g] = \E[\Delta Y_{t}(0) | U=1]
\end{align*}

:::

<br>

. . .

Using very similar arguments as before, can show that 
\begin{align*}
  ATT(g,t) = \E[Y_t - Y_{g-1} | G=g] - \E[Y_t - Y_{g-1} | U=1]
\end{align*}

. . .

where the main difference is that we use $(g-1)$ as the "base period" (this is the period right before group $g$ becomes treated).



## Extensions/Summary

The previous discussion emphasizes a general purpose identification strategy with staggered treatment adoption:

. . .

Step 1: Target disaggregated treatment effect parameters (i.e., group-time average treatment effects)

* You can use many existing approaches for this step (generally with very minor modification) that work for smaller problems without staggered treatment adoption

* Discussion above has been for DID identification strategy, but other possibilities fit into this framework: conditioning on lagged outcomes, unit-specific linear trends, interactive fixed effects, change-in-changes, triple differences, etc.

. . .

Step 2: (If desired) combine disaggregated treatment effects into lower dimensional summary treatment effect parameter


## Estimation

CS

Other approaches

## Other Issues

Selecting comparison group

anticipation

## Limitation of TWFE Regressions in this Context

## Empirical Example: Minimum Wages and Employment

- Use county-level data from 2003-2007 during a period where the federal minimum wage was flat

. . . 

- Exploit minimum wage changes across states

    - Any state that increases their minimum wage above the federal minimum wage will be considered as treated
  
. . . 

- Interested in the effect of the minimum wage on teen employment

. . . 

- We'll also make a number of simplifications:
  * not worry much about issues like clustered standard errors
  * not worry about variation in the amount of the minimum wage change (or whether it keeps changing) across states

. . . 

[Goals: ]{.alert}

* Get some experience with an application and DID-related code

* Assess how much do the issues that we have been talking about matter in practice



## Code

Full code is available on my website: [https://bcallaway11.github.io/files/presentations/northwestern-causal-inference-workshop](https://bcallaway11.github.io/files/presentations/northwestern-causal-inference-workshop) or link is on my homepage [brantlycallaway.com](https://www.brantlycallaway.com)

<span class="alert">R packages used in empirical example</span>

```{r, message=FALSE}
library(did)
library(BMisc)
library(twfeweights)
library(fixest)
library(modelsummary)
library(ggplot2)
load(url("https://github.com/bcallaway11/did_chapter/raw/master/mw_data_ch2.RData"))
```



## Setup Data

```{r}
# drops NE region and a couple of small groups
mw_data_ch2 <- subset(mw_data_ch2, (G %in% c(2004,2006,2007,0)) & (region != "1"))
head(mw_data_ch2[,c("id","year","G","lemp","lpop","region")])
# drop 2007 as these are right before fed. minimum wage change
data2 <- subset(mw_data_ch2, G!=2007 & year >= 2003)
# keep 2007 => larger sample size
data3 <- subset(mw_data_ch2, year >= 2003)
```


## TWFE Regression

```{r}
twfe_res2 <- fixest::feols(lemp ~ post | id + year,
                           data=data2,
                           cluster="id")
```

<br>

```{r}
modelsummary(list(twfe_res2), gof_omit=".*")
```




## $ATT(g,t)$ (Callaway and Sant'Anna)

```{r warning=FALSE}
attgt <- did::att_gt(yname="lemp",
                     idname="id",
                     gname="G",
                     tname="year",
                     data=data2,
                     control_group="nevertreated",
                     base_period="universal")
tidy(attgt)[,1:5] # print results, drop some extra columns
```





## Plot $ATT(g,t)$'s


```{r, fig.align="center", fig.width=10, fig.height=8, echo=FALSE}
ggdid(attgt)
```



## Compute $ATT^O$

```{r}
attO <- did::aggte(attgt, type="group")
summary(attO)
```




## Comments

The differences between the CS estimates and the TWFE estimates are fairly large here: the CS estimate is about 50% larger than the TWFE estimate, though results are qualitatively similar.

. . . 

[Additional Details](DID-Introduction.html#/de-chaisemartin-and-dhaultfoeuille-weights)

## de Chaisemartin and d'Haultfoeuille weights {visibility="uncounted"}

```{r echo=FALSE, fig.align="center", fig.width=10, fig.height=8}
tw <- twfeweights::twfe_weights(attgt)
tw <- tw[tw$G != 0,]
tw$post <- as.factor(1*(tw$TP >= tw$G))
twfe_est <- sum(tw$wTWFEgt*tw$attgt)
ggplot(data=tw,
       mapping=aes(x=wTWFEgt, y=attgt, color=post)) +
  geom_hline(yintercept=0, linewidth=1.5) +
  geom_vline(xintercept=0, linewidth=1.5) + 
  geom_point(size=6) +
  theme_bw() +
  ylim(c(-.15,.05)) + xlim(c(-.4,.7))
```



## $ATT^O$ weights {visibility="uncounted"}

```{r echo=FALSE, fig.align="center", fig.width=10, fig.height=8}
wO <- attO_weights(attgt)
wO <- wO[wO$G != 0,]
attO_est = sum(wO$wOgt*wO$attgt)
wO$post <- as.factor(1*(wO$TP >= wO$G))
ggplot(data=wO,
       mapping=aes(x=wOgt, y=attgt, color=post)) +
  geom_hline(yintercept=0, linewidth=1.5) +
  geom_vline(xintercept=0, linewidth=1.5) + 
  geom_point(shape=18, size=8) +
  theme_bw() +
  ylim(c(-.15,.05)) + xlim(c(-.4,.7))
```



## Weight Comparison {visibility="uncounted"}

```{r echo=FALSE, fig.align="center", fig.width=10, fig.height=8}
plot_df <- cbind.data.frame(tw, wOgt=wO$wOgt)
plot_df <- plot_df[plot_df$post==1,]
plot_df$g.t <- as.factor(paste0(plot_df$G,",",plot_df$TP))

ggplot(plot_df, aes(x=wTWFEgt, y=attgt, color=g.t)) +
  geom_point(size=6) +
  theme_bw() +
  ylim(c(-.15,.05)) + xlim(c(-.4,.7)) + 
  geom_point(aes(x=wOgt), shape=18, size=8) +
  geom_hline(yintercept=0, linewidth=1.5) +
  geom_vline(xintercept=0, linewidth=1.5) + 
  xlab("weight")
```



## Discussion {visibility="uncounted"}

To summarize: $ATT^O = -0.057$ while $\alpha^{TWFE} = -0.038$.  This difference can be fully accounted for

* Pre-treatment differences in paths of outcomes across groups: explains about 64% of the difference

* Differences in weights applied to the same post-treatment $ATT(g,t)$: explains about 36% of the difference. [If you apply the post-treatment weights and "zero out" pre-treatment differences, the estimate would be $-0.050$.]

```{r echo=FALSE, eval=FALSE}
twfe_post <- sum(tw$wTWFEgt[tw$post==1] * tw$attgt[tw$post==1])
twfe_post

# pre-treatment contamination/bias
pre_bias <- sum(tw$wTWFEgt[tw$post==0] * tw$attgt[tw$post==0])
pre_bias

twfe_bias <- twfe_est - attO_est
pre_bias/twfe_bias # bias from pre-treatment PTA violations
(twfe_post-attO_est)/twfe_bias # bias from TWFE weights instead of ATT^O weights
```

. . . 

In my experience: this is fairly representative of how much new DID approaches matter relative to TWFE regressions.  It does not seem like "catastrophic failure" of TWFE, but (in my view) these are meaningful differences (and, e.g., given slightly different $ATT(g,t)$'s, the difference in the weighting schemes could change the qualitative results).

* Of course, this whole discussion hinges crucially on how much treatment effect heterogeneity there is.  More TE Het $\implies$ more sensitivity to weighting schemes [just looking at TWFE regression does not give insight into how much TE Het there is.]



## Additional Comments on Weights {visibility="uncounted"}

One more comment: there is a lot concern about [negative weights]{.alert} (both in econometrics and empirical work).  

. . .

* There were no negative weights in the example above, but the weights still weren't great.  

    * No negative weights does rule out "sign reversal"

. . .

* But, in my view, the [more important issue is the non-transparent weighting scheme]{.alert}.
    * Example 1: If you try using `data3` (the data that includes $G=2007$), you will get a negative weight on $ATT(g=2004,t=2007)$.  But it turns out not to matter much, and TWFE works better in this case than in the case that I showed you.
    * Example 2: Alternative treatment effect parameter $\rightarrow$



## "Simple" Aggregation {visibility="uncounted"}

Consider the following alternative aggregated treatment effect parameter
\begin{align*}
  ATT^{simple} := \sum_{t=g}^\mathcal{T} ATT(g,t) \frac{\P(G=g | G \in \bar{\mathcal{G}})}{\sum_{t=g}^{\mathcal{T}} \P(G=g| G \in \bar{\mathcal{G}})}
\end{align*}
Consider imputation so that you have $Y_{it}-\hat{Y}_{it}(0)$ available in all periods.  This is the $ATT$ parameter that you get by averaging all of those.

. . . 

Relative to $ATT^O$, early treated units get more weight (because we have more $Y_{it}-\hat{Y}_{it}(0)$ for them).

. . . 

By construction, weights are all positive.  However, they are different from $ATT^O$ weights



## "Simple" Aggregation {visibility="uncounted"}

```{r echo=FALSE, eval=FALSE}
att_simple_est <- aggte(attgt, type="simple")$overall.att #-0.0646
(att_simple_est - attO_est) / attO_est

```{r echo=FALSE, fig.align="center", fig.width=10, fig.height=8}
# calculate att^simple weights
wsimple <- att_simple_weights(attgt)
wsimple <- wsimple[wsimple$G != 0,]
attsimple_est <- sum(wsimple$wsimplegt*wsimple$attgt)
wsimple$post <- as.factor(1*(wsimple$TP >= wsimple$G))

# comparison of att^O and att^simple weights
plot_df <- cbind.data.frame(wsimple, wOgt=wO$wOgt)
plot_df <- plot_df[plot_df$post==1,]
plot_df$g.t <- as.factor(paste0(plot_df$G,",",plot_df$TP))

ggplot(plot_df, aes(x=wsimplegt, y=attgt, color=g.t)) +
  geom_point(shape=15, size=6) +
  theme_bw() +
  ylim(c(-.15,.05)) + xlim(c(-.4,.7)) + 
  geom_point(aes(x=wOgt), shape=18, size=8) +
  geom_hline(yintercept=0, linewidth=1.5) +
  geom_vline(xintercept=0, linewidth=1.5) + 
  xlab("weight")
```



## "Simple" Aggregation {visibility="uncounted"}

Besides the violations of parallel trends in pre-treatment periods, these weights are further away from $ATT^O$ than the TWFE regression weights are!

. . . 

In fact, you calculate $ATT^{simple} = -0.065$ (13% larger in magnitude that $ATT^O$)

. . . 

Finally, if you are "content with" non-negative weights, then you can get any summary measure from $-0.019$ (the smallest $ATT(g,t)$) to $-0.13$ (the largest).  This is a wide range of estimates.

. . . 

In my view, the discussion above suggests that clearly stating a target aggregate treatment effect parameter and choosing weights that target that parameter is probably more important than checking for negative weights
